Learning for Logistic Regression Model in Parallel

### 2013-06-24 draft.tex
version after proofread

\begin{abstract}
Logistic regression (LR) model has various applications in machine learning.
It is a supervised learning model for pattern classification and enjoys the benefits of statistical principles.
This paper addresses the issue of computational efficiency for solving LR in the big data scenario.
We compare the performance between different computing platforms and different parallel algorithms.
Our goal is to provide valuable suggestions on choosing computing platforms and learning algorithms according to different application situations for specific datasets and machine resources.
Three different parallel computing platforms along with three different types of parallel algorithms are carefully studied in this paper.
In addition, we present a novel parallel sublinear method to optimize the solution of LR.
%Hadoop, Mahout and Spark are three existing parallel computing platforms that can be implemented with large scale machine learning algorithms.
%Parallel gradient Descent and stochastic gradient descent are two state-of-art algorithms for solving LR.
To enhance capability, we consider the tradeoff between efficiency and accuracy. Extensive experimental results verify the effective performance of proposed algorithms.
\end{abstract}

@@

In recent years, many modern datasets are increasingly larger and larger.
A most identifiable characteristic of these datasets is that data scale is very large and data dimensionality is very high.
Large data volume and high data dimensionality bring unavoidable computational challenges to machine learning problems.
For example, in social networks, we have datasets~\cite{lewis2008tastes} covering multiple dimensions like friendship relationship and share of common interests. They generally include millions of records over millions of attributes. More evidently, for the task of text categorization, we usually have to handle datasets with data volumes in billion scales and each data instance is characterized by thousands of feature dimensions~\cite{lewis2004rcv1}.
We tackle these challenges on both algorithm level and computing platform level via parallelization methods.
For many applications, in order to attain speedup in performance, we can employ not only task parallelism, but also data parallelism~\cite{subhlok1993exploiting}.
In this paper, we mainly discuss issues in task parallelism.

@@

\subsection{Scalability Solutions}
%
In big data scenarios, large datasets require us to develop machine learning algorithms towards a more efficient and more parallelized end.
Researchers have already done much work to cater for the needs of massive datasets.
Early work like Parallel Support Vector Machines (PSVM)~\cite{chang2011psvm} employed an approximation matrix decomposition method based on row selections to reduce memory usage.
It can then increase the number of parallel computational nodes to several hundreds.
Later, the work of Parallel Latent Dirichlet Allocation (PLDA)~\cite{wang2009plda} improves the computational efficiency of the LDA model by means of sampling methods.
The parallel algorithm in Hadoop is robust, for it has fault tolerance of machine failures, thus taking advantage of Hadoop features.
Recently, Dean \etal~\cite{le2011building} claimed that the advantage of parallelization is fully taken in deep learning algorithms.
It pushed the limit of parallel computational nodes to a hundred-million level, and in the meantime, achieved the best learning performance ever.
GraphLab~\cite{kyrola2012graphchi} is a recently developed tool for large-scale machine learning. It tackles efficiency and scalability problems of machine leaning algorithms by graph-theoretical models.

\subsection{Computing Platforms}\label{sec:platform}
Three computing platforms that we are working on have their own advantages to implement machine learning algorithms.
The Apache Hadoop~\cite{white2012hadoop} software library allows for distributed processing of large datasets across clusters of computers using simple programming models.
Hadoop utilizes MapReduce~\cite{dean2008mapreduce} as its computational paradigm.
MapReduce paradigm takes a set of key-value pairs as its input.
Then a user-defined \textit{map} function turns the input into some intermediate key-value pairs.
Another user-defined \textit{reduce} function merges values for the same intermediate pairs to produce output.
All of the operations are independent, thereby making MapReduce paradigm easy to be parallelized.
In addition, Hadoop provides a Distributed File System (HDFS).
Both MapReduce and HDFS are designed to handle node failures in an automatic way, making Hadoop support large clusters that are built on commodity hardware.
The goal of Apache Mahout~\cite{mahoutscalable} is to build scalable machine learning libraries. Their core algorithms for clustering, classification and collaborative filtering are implemented on top of Apache Hadoop using the MapReduce paradigm. The core libraries are highly optimized so that Apache Mahout also has good performance for non-distributed algorithms.
Spark~\cite{zaharia2010spark} is a cluster computing system that aims to make data analysis fast.
Spark supports in-memory cluster computing. A job can load data into memory and query it repeatedly by creating and caching resilient distributed datasets (RDDs).
Moreover, RDDs achieve fault tolerance through lineage: information about how RDDs are derived from other RDDs is stored in reliable storage, thus making RDDs easy to be rebuilt if a certain partition is lost.
Spark can run up to 100 times faster than Hadoop for iterative algorithms.

In Table~\ref{tab:systemcmp}, we compare different features between Hadoop and Spark computing platforms.
As Mahout is mainly built on Hadoop, it inherits most of Hadoop features.
For simplicity, we use short abbreviations in the table.
Spark supports two types of operations on RDDs: Actions and Transformations (As and Ts).
Actions include functions like \textit{count}, \textit{collect} and \textit{save}. They usually return a result from input RDDs.
Transformations include functions like \textit{map}, \textit{filter} and \textit{join}. They normally build new RDDs from other RDDs, which comply with the lineage rule.
In the table, we use "NFS" to represent "Normal File System" and "Fault Tolerance" here is specifically for node failure situations.
\begin{table}[h]
\centering
\caption{Platform Comparison: Hadoop vs. Spark}\label{tab:systemcmp}\vspace{-0.3cm}
\begin{tabular}{|c|c|c|}
\hline
           & Hadoop & Spark \\
\hline
Computational Paradigm  & MapReduce & As and Ts \\
\hline
File System Supported & HDFS & HDFS and NFS \\
\hline
Design Concept  & Key-value Pairs & RDD \\
\hline
Fault Tolerance Technique & Redundancy & Lineage \\
\hline
\end{tabular}\vspace{-0.5cm}
\end{table}

\subsection{Sublinear Methods}
Sublinear methods are recently proposed by many researchers.
Clarkson \etal~\cite{clarkson2010sublinear} first presented the solution of approximation algorithms in sublinear time.
It is applied to optimization problems for linear classifiers and minimum enclosing balls.
The algorithm employs a novel sampling technique along with a new multiplicative update procedure.
Hazan \etal~\cite{hazanbeating} exploited the approach to linear SVM.
Cotter \etal~\cite{cotter2012kernelized} extended the method to kernelized SVM.
In~\cite{hazan2011optimal}, Hazan \etal applied the sublinear approximation approach to solve linear regression with penalties.
Garber and Hazan~\cite{garberapproximating} also developed it in Semi-Definite Programming (SDP).
Peng \etal~\cite{peng2012sublinear} utilized the method in the LR model with penalties and developed sequential sublinear algorithms for both $\ell_1$-penalty and $\ell_2$-penalty.

@@

\section{Experimental Setup} \label{sec:setup}
In this section, we present the details of the dataset information, cluster information, and the test programs of our experiments.

@@

\section{Experimental Results} \label{sec:experiment}
In this section, we conduct the empirical analysis.

@@

\subsection{Results on Precision} \label{sec:precision}
The results of six test programs achieved on five datasets are shown in Table~\ref{tab:table3}.
%
\begin{table}[h]
\centering
\caption{Accuracy Results. The meanings of abbreviations are as follows: 20-N-G, 20 News Group; URL-R, URL-Reputation.}\label{tab:table3}\vspace{-0.3cm}
\begin{tabular}{|c|c|c|c|c|}
\hline
           & 20-N-G & Gisette & ECUESpam & URL-R \\
\hline
Mahout     & 71.3\% & 91.5\% & 85.2\% & 91.5\% \\
\hline
Liblinear  & 92.0\% & 97.4\% & 97.1\% & --- \\
\hline
SLLR       & 91.5\% & 94.8\% & 92.3\% & 94.2\% \\
\hline
PSUBPLR-MR & 90.5\% & 94.6\% & 91.7\% & 93.8\% \\
\hline
PGDPLR-SPARK & 92.0\% & 97.0\% & 93.7\% & 96.0\% \\
\hline
PSUBPLR-SPARK & 90.5\% & 95.8\% & 91.7\% & 94.0\% \\
\hline
\end{tabular}
\end{table}
%
These are the averaged results of cross validation.
In Figures~\ref{fig:accuracy} (a)-(d), we show the test error as a function of iteration number on each dataset for all six test programs.
Note that Liblinear cannot be implemented with the full \textbf{URL-Reputation} dataset on our machine due to memory limit.
This proves the scalability issue of Liblinear.
In the tests, Liblinear can handle $33.3\%$ of the whole data volume at most, with the test accuracy of $96.2\%$ and the running time of 519s.
%
\begin{figure*}[tb]
\begin{center}
\begin{tabular}{cccc}
   %\includegraphics[height=4.5cm,width=5cm]{img/2d_accuracy_iteration.eps}&
   \includegraphics[height=4.cm,width=4.5cm]{img/20NewsGroup_accuracy_iteration.eps}&
   \hspace{-0.6cm}\includegraphics[height=4.cm,width=4.5cm]{img/Gisette_accuracy_iteration.eps}&
   \hspace{-0.6cm}\includegraphics[height=4.cm,width=4.5cm]{img/ECUESpam_accuracy_iteration.eps}&
   \hspace{-0.6cm}\includegraphics[height=4.cm,width=4.5cm]{img/URL-Reputation_accuracy_iteration.eps}\\
   (a) \textbf{20NewsGroup}. & \hspace{-0.3cm}(b) \textbf{Gisette}. & \hspace{-0.3cm}(c) \textbf{ECUESpam}. & \hspace{-0.3cm}(d) \textbf{URL-Reputation}.\\
   \end{tabular}
\end{center}\vspace{-0.3cm}
   \caption{Test error, as a function of iteration number.}\vspace{-0.5cm}
\label{fig:accuracy}
\end{figure*}

@@

The running time of six test programs used on five datasets is shown in Table~\ref{tab:table4}.
These are the averaged results corresponding to the accuracies we achieved in Table~\ref{tab:table3}.
%
\begin{table}[h]
\centering
\caption{Running Time. The meanings of abbreviations are as follows: 20-N-G, 20 News Group; URL-R, URL-Reputation.}\label{tab:table4}\vspace{-0.3cm}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
           & 20-N-G & Gisette & ECUESpam & URL-R \\
\hline
Mahout     & 9.83s & 131.8s & 96s & 10100s \\
\hline
Liblinear  & 0.79s & 2.4s & 13s & --- \\
\hline
SLLR       & 20.05s & 130.5s & 1028s & 3248s \\
\hline
PSUBPLR-MR & 1360.85s & 3687.9s & 11478s & 16098s \\
\hline
PGDPLR-SPARK & 10.52s & 99.2s & 924s & 3615s \\
\hline
PSUBPLR-SPARK & 8.57s & 89.1s & 796s & 2918s \\
\hline
\end{tabular}
\end{table}

@@

5) Spark employs the ``all-in-memory" strategy, and it constructs RDD on demand. We implement PGDPLR-SPARK and PSUBPLR-SPARK in normal file system instead of HDFS.
          Current results on Spark show that it is much more efficient than Hadoop, and performs better than Mahout.
          In the parallel situation, we recommend Spark to process massive data. Further, we recommend to choose PSUBPLR-SPARK for less running time in exchange for a little bit precision loss.
          As the platform is still under developing process, we expect that our running time results for PGDPLR-SPARK and PSUBPLR-SPARK can still be improved.
          For example, Spark currently only supports for one reducer, thus it does not use full CPU power when performing reduce.

@@

\subsection{Results on Changing Cluster}
In Figures~\ref{fig:time} (a)-(d), we show the running time as a function of used node number on each dataset for all three parallel test programs.
%
\begin{figure*}[tb]
\begin{center}
\begin{tabular}{cccc}
   %\includegraphics[height=4.5cm,width=5cm]{img/2d_time.eps}&
   \includegraphics[height=4.cm,width=4.5cm]{img/20NewsGroup_time.eps}&
   \hspace{-0.6cm}\includegraphics[height=4.cm,width=4.5cm]{img/Gisette_time.eps}&
   \hspace{-0.6cm}\includegraphics[height=4.cm,width=4.5cm]{img/ECUESpam_time.eps}&
   \hspace{-0.6cm}\includegraphics[height=4.cm,width=4.5cm]{img/URL-Reputation_time.eps}\\
   (a) \textbf{20NewsGroup}. & \hspace{-0.3cm}(b) \textbf{Gisette}. & \hspace{-0.3cm}(c) \textbf{ECUESpam}. & \hspace{-0.3cm}(d) \textbf{URL-Reputation}.\\
   \end{tabular}
\end{center}\vspace{-0.3cm}
   \caption{Running time, as a function of used node number.}\vspace{-0.5cm}
\label{fig:time}
\end{figure*}
%
As we can see from the results, fewer computation resources provided, longer the implementation process will take.
However, this trend is not in a consistent fashion. When the number of machines gets higher, it will make the less impact.
We may reach a point for every algorithm when running time halts to decrease as machine number increases.
This shows the true scalability of the designed algorithm, which we can study in our future work.

@@

We show the iteration time as a function of percentage of failed maps on \textbf{URL-Reputation} dataset for all three parallel test programs in Fig.~\ref{fig:14}.
This is tested on 6 nodes.
\begin{figure}[tb]
\center \includegraphics[height=4.5cm,width=9cm]{img/fault_tolerance.eps}\vspace{-0.3cm}
\caption{Iteration time, as a function of percentage of failed maps, on \textbf{URL-Reputation} Dataset, run on 6 nodes}\label{fig:14}\vspace{-0.5cm}
\end{figure}
As Fig.~\ref{fig:14} shows, iteration time of PSUBPLR-MR remains unaffected when maps fail. But it takes the cost of precision loss if we enforce that no additional iterations are implemented.
PGDPLR-SPARK and PSUBPLR-SPARK will increase iteration time when the maps fail. It is due to RDD reconstruction. The increase is not significant.
This difference also reveals the different mechanism of fault tolerance between Hadoop and Spark.

\section{Conclusion} \label{sec:concl}
In this paper we analyzed three optimization approaches along with three computing platforms to train LR model on large-scale, high-dimensional datasets for classification. In addition, we presented a novel parallel sublinear algorithm for LR implemented on both Hadoop and Spark. Based on extensive experiments, we summarized key features of each algorithm implemented on different computing platforms.
Our future work will focus on combining algorithm design and computing platforms to optimize classifiers in machine learning.

>>>>

\begin{abstract}
Logistic regression (LR) model has various applications in machine learning.
LR is a supervised learning model for pattern classification and enjoys the benefits of statistical principles.
This paper addresses the issue of computational efficiency for solving LR in the big data scenario.
We compare the performance between different computing platforms and different parallel algorithms.
Our goal is to provide valuable suggestions on choosing computing platforms and learning algorithms according to different application situations for specific datasets and machine resources.
Three different parallel computing platforms along with three different types of parallel algorithms are carefully studied in this paper.
In addition, we present a novel parallel sublinear method to optimize the solution of LR.
%Hadoop, Mahout and Spark are three existing parallel computing platforms that can be implemented with large scale machine learning algorithms.
%Parallel gradient Descent and stochastic gradient descent are two state-of-art algorithms for solving LR.
To enhance capability, we consider the tradeoff between efficiency and accuracy. Extensive experimental results verify the effective performance of proposed algorithms.
\end{abstract}

@@

In recent years, many modern datasets grow drastically in both data volume and data dimensionality.
Large data volume and high data dimensionality bring unavoidable computational challenges to machine learning problems.
For example, in social networks, we have datasets~\cite{lewis2008tastes} covering multiple dimensions like friendship relationship and share of common interests. They generally include millions of records over millions of attributes. More evidently, for the task of text and multimedia categorization, we usually have to handle datasets with data volumes in billion scales and each data instance is characterized by thousands of feature dimensions~\cite{chang2011foundations}.
We tackle these challenges on both algorithm level and computing platform level via parallelization methods.
For many applications, in order to attain speedup in performance, we can employ not only task parallelism, but also data parallelism~\cite{subhlok1993exploiting}.
In this paper, we mainly discuss issues in task parallelism.

@@

\subsection{Scalability Solutions}
%
In big data scenarios, large datasets require us to develop machine learning algorithms towards a more efficient and more parallelized end.
Researchers have already done much work to cater for the needs of massive datasets.
Early work like Parallel Support Vector Machines (PSVM)~\cite{chang2011psvm} employed an approximation matrix decomposition method based on row selections to reduce memory usage.
It can then increase the number of parallel computational nodes to several hundreds.
Later, the work of Parallel Latent Dirichlet Allocation (PLDA)~\cite{wang2009plda}~\cite{newman2007distributed} improves the computational efficiency of the LDA model by means of sampling methods.
The parallel algorithm in Hadoop is robust, for it has fault tolerance of machine failures, thus taking advantage of Hadoop features.
Recently, Dean \etal~\cite{le2011building} claimed that the advantage of parallelization is fully taken in deep learning algorithms.
It pushed the limit of parallel computational nodes to a hundred-million level, and in the meantime, achieved the best learning performance ever.
GraphLab~\cite{kyrola2012graphchi} is a recently developed tool for large-scale machine learning. It tackles efficiency and scalability problems of machine leaning algorithms by graph-theoretical models.

\subsection{Computing Platforms}\label{sec:platform}
In this paper, we specifically compares parallelization efforts of LR in two dimensions: platform and algorithm.

Three computing platforms that we are working on have their own advantages to implement machine learning algorithms.
The Apache Hadoop~\cite{white2012hadoop} software library allows for distributed processing of large datasets across clusters of computers using simple programming models.
Hadoop utilizes MapReduce~\cite{dean2008mapreduce} as its computational paradigm.
MapReduce paradigm takes a set of key-value pairs as its input.
Then a user-defined \textit{map} function turns the input into some intermediate key-value pairs.
Another user-defined \textit{reduce} function merges values for the same intermediate pairs to produce output.
All of the operations are independent, thereby making MapReduce paradigm easy to be parallelized.
In addition, Hadoop provides a Distributed File System (HDFS).
Both MapReduce and HDFS are designed to handle node failures in an automatic way, making Hadoop support large clusters that are built on commodity hardware.
The goal of Apache Mahout~\cite{mahoutscalable} is to build scalable machine learning libraries. Their core algorithms for clustering, classification and collaborative filtering are implemented on top of Apache Hadoop using the MapReduce paradigm. The core libraries are highly optimized so that Apache Mahout also has good performance for non-distributed algorithms.
Spark~\cite{zaharia2010spark} is a cluster computing system that aims to make data analysis fast.
Spark supports in-memory cluster computing. A job can load data into memory and query it repeatedly by creating and caching resilient distributed datasets (RDDs).
Moreover, RDDs achieve fault tolerance through lineage: information about how RDDs are derived from other RDDs is stored in reliable storage, thus making RDDs easy to be rebuilt if a certain partition is lost.
Spark can run up to 100 times faster than Hadoop for iterative algorithms.

In Table~\ref{tab:systemcmp}, we compare different features between Hadoop and Spark computing platforms.
As Mahout is mainly built on Hadoop, it inherits most of Hadoop features.
For simplicity, we use short abbreviations in the table.
Spark supports two types of operations on RDDs: Actions and Transformations (As and Ts).
Actions include functions like \textit{count}, \textit{collect} and \textit{save}. They usually return a result from input RDDs.
Transformations include functions like \textit{map}, \textit{filter} and \textit{join}. They normally build new RDDs from other RDDs, which comply with the lineage rule.
In the table, we use ``NFS" to represent ``Normal File System" and "Fault Tolerance" here is specifically for node failure situations.
\begin{table}[h]
\centering
\caption{Platform Comparison: Hadoop vs. Spark}\label{tab:systemcmp}\vspace{-0.3cm}
\begin{tabular}{|c|c|c|}
\hline
           & Hadoop & Spark \\
\hline
Computational Paradigm  & MapReduce & As and Ts \\
\hline
File System Supported & HDFS & HDFS and NFS \\
\hline
Design Concept  & Key-value Pairs & RDD \\
\hline
Fault Tolerance Technique & Redundancy & Lineage \\
\hline
\end{tabular}\vspace{-0.5cm}
\end{table}

\subsection{Sublinear Methods}
Sublinear methods are recently proposed by many researchers.
Clarkson \etal~\cite{clarkson2010sublinear} first presented the solution of approximation algorithms in sublinear time.
It is applied to optimization problems for linear classifiers and minimum enclosing balls.
The algorithm employs a novel sampling technique along with a new multiplicative update procedure.
Hazan \etal~\cite{hazanbeating} exploited the approach to linear SVM.
Cotter \etal~\cite{cotter2012kernelized} extended the method to kernelized SVM.
%In~\cite{hazan2011optimal}, Hazan \etal applied the sublinear approximation approach to solve linear regression with penalties.
Garber and Hazan~\cite{garberapproximating} also developed it in Semi-Definite Programming (SDP).
Peng \etal~\cite{peng2012sublinear} utilized the method in the LR model with penalties and developed sequential sublinear algorithms for both $\ell_1$-penalty and $\ell_2$-penalty.

@@

\section{Experimental Setup} \label{sec:setup}
This section presents the details of the dataset information, cluster information, and the test programs of our experiments.

@@

\section{Experimental Results} \label{sec:experiment}
In this section, we conduct empirical analysis on experimental results.
We show three things: 1) correctness of our test program, 2) accuracy comparison, and 3) speedup comparison.

@@

\subsection{Results on Precision} \label{sec:precision}
The results of six test programs achieved on five datasets are shown in Table~\ref{tab:table3}.
%
\begin{table}[h]
\centering
\caption{Accuracy Results. The meanings of abbreviations are as follows: 20-N-G, 20 News Group; URL-R, URL-Reputation.}\label{tab:table3}\vspace{-0.3cm}
\begin{tabular}{|c|c|c|c|c|}
\hline
           & 20-N-G & Gisette & ECUESpam & URL-R \\
\hline
Mahout     & 71.3\% & 91.5\% & 85.2\% & 91.5\% \\
\hline
Liblinear  & \textbf{92.0\%} & \textbf{97.4\%} & \textbf{97.1\%} & --- \\
\hline
SLLR       & 91.5\% & 94.8\% & 92.3\% & 94.2\% \\
\hline
PSUBPLR-MR & 90.5\% & 94.6\ & 91.7\% & 93.8\% \\
\hline
PGDPLR-SPARK & \textbf{92.0\%} & 97.0\% & 93.7\% & \textbf{96.0\%} \\
\hline
PSUBPLR-SPARK & 90.5\% & 95.8\% & 91.7\% & 94.0\% \\
\hline
\end{tabular}
\end{table}
%
These are the averaged results of cross validation.
In Figures~\ref{fig:accuracy} (a)-(d), we show the test error as a function of iteration number on each dataset for all six test programs.
Note that Liblinear cannot be implemented with the full \textbf{URL-Reputation} dataset on our machine due to memory limit.
This proves the scalability issue of Liblinear.
In the tests, Liblinear can handle $33.3\%$ of the whole data volume at most, with the test accuracy of $96.2\%$ and the running time of 519s.
%
\begin{figure*}[tb]
\begin{center}
\begin{tabular}{cccc}
   %\includegraphics[height=4.5cm,width=5cm]{img/2d_accuracy_iteration.eps}&
   \includegraphics[height=4.cm,width=4.5cm]{img/20NewsGroup_accuracy_iteration.eps}&
   \hspace{-0.6cm}\includegraphics[height=4.cm,width=4.5cm]{img/Gisette_accuracy_iteration.eps}&
   \hspace{-0.6cm}\includegraphics[height=4.cm,width=4.5cm]{img/ECUESpam_accuracy_iteration.eps}&
   \hspace{-0.6cm}\includegraphics[height=4.cm,width=4.5cm]{img/URL-Reputation_accuracy_iteration.eps}\\
   (a) \textbf{20NewsGroup}. & \hspace{-0.3cm}(b) \textbf{Gisette}. & \hspace{-0.3cm}(c) \textbf{ECUESpam}. & \hspace{-0.3cm}(d) \textbf{URL-Reputation}.\\
   \end{tabular}
\end{center}\vspace{-0.3cm}
   \caption{Test error, as a function of iteration number.}\vspace{-0.5cm}
\label{fig:accuracy}
\end{figure*}

@@

The running time of six test programs used on five datasets is shown in Table~\ref{tab:table4}.
These are the averaged results corresponding to the accuracies we achieved in Table~\ref{tab:table3}.
%
\begin{table}[h]
\centering
\caption{Running Time. The meanings of abbreviations are as follows: 20-N-G, 20 News Group; URL-R, URL-Reputation.}\label{tab:table4}\vspace{-0.3cm}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
           & 20-N-G & Gisette & ECUESpam & URL-R \\
\hline
Mahout     & 9.83s & 131.8s & 96s & 10100s \\
\hline
Liblinear  & \textbf{0.79s} & \textbf{2.4s} & \textbf{13s} & --- \\
\hline
SLLR       & 20.05s & 130.5s & 1028s & 3248s \\
\hline
PSUBPLR-MR & 1360.85s & 3687.9s & 11478s & 16098s \\
\hline
PGDPLR-SPARK & 10.52s & 99.2s & 924s & 3615s \\
\hline
PSUBPLR-SPARK & 8.57s & 89.1s & 796s & \textbf{2918s} \\
\hline
\end{tabular}
\end{table}

@@

5) Spark employs the ``all-in-memory" strategy, and it constructs RDD on demand. We implement PGDPLR-SPARK and PSUBPLR-SPARK in normal file system instead of HDFS.
          Current results on Spark show that it is much more efficient than Hadoop, and performs better than Mahout.
          In the parallel situation, we recommend Spark to process massive data. Further, we recommend to choose PSUBPLR-SPARK for less running time in exchange for a little bit precision loss.
          As the platform is still under developing process, we expect that our running time results for PGDPLR-SPARK and PSUBPLR-SPARK can still be improved.

@@

\subsection{Results on Changing Cluster}
In Figures~\ref{fig:time} (a)-(d), we show the running time as a function of used node number on each dataset for all three parallel test programs.
%
\begin{figure*}[tb]
\begin{center}
\begin{tabular}{cccc}
   %\includegraphics[height=4.5cm,width=5cm]{img/2d_time.eps}&
   \includegraphics[height=3.9cm,width=4.5cm]{img/20NewsGroup_time.eps}&
   \hspace{-0.6cm}\includegraphics[height=3.9cm,width=4.5cm]{img/Gisette_time.eps}&
   \hspace{-0.6cm}\includegraphics[height=3.9cm,width=4.5cm]{img/ECUESpam_time.eps}&
   \hspace{-0.6cm}\includegraphics[height=3.9cm,width=4.5cm]{img/URL-Reputation_time.eps}\\
   (a) \textbf{20NewsGroup}. & \hspace{-0.3cm}(b) \textbf{Gisette}. & \hspace{-0.3cm}(c) \textbf{ECUESpam}. & \hspace{-0.3cm}(d) \textbf{URL-Reputation}.\\
   \end{tabular}
\end{center}\vspace{-0.3cm}
   \caption{Running time, as a function of used node number.}\vspace{-0.5cm}
\label{fig:time}
\end{figure*}
%
As we can see from the results, fewer computation resources provided, longer the implementation process will take.
However, this trend is not in a consistent fashion. When the number of machines gets higher, it will make the less impact.
We may reach a point for every algorithm when running time halts to decrease as machine number increases.
This shows the true scalability of the designed algorithm, which we can study in our future work.

@@

We show the iteration time as a function of percentage of failed maps on \textbf{URL-Reputation} dataset for all three parallel test programs in Fig.~\ref{fig:14}.
This is tested on 6 nodes.
\begin{figure}[tb]
\center \includegraphics[height=4cm,width=8cm]{img/fault_tolerance.eps}\vspace{-0.3cm}
\caption{Iteration time, as a function of percentage of failed maps, on \textbf{URL-Reputation} Dataset, run on 6 nodes}\label{fig:14}\vspace{-0.5cm}
\end{figure}
As Fig.~\ref{fig:14} shows, iteration time of PSUBPLR-MR remains unaffected when maps fail. But it takes the cost of precision loss if we enforce that no additional iterations are implemented.
PGDPLR-SPARK and PSUBPLR-SPARK will increase iteration time when the maps fail. It is due to RDD reconstruction. The increase is not significant.
This difference also reveals the different mechanism of fault tolerance between Hadoop and Spark.

\section{Conclusion} \label{sec:concl}
In this paper we analyzed three optimization approaches along with three computing platforms to train LR model on large-scale, high-dimensional datasets for classification. In addition, we presented a novel parallel sublinear algorithm for LR implemented on both Hadoop and Spark. Based on extensive experiments, we summarized key features of each algorithm implemented on different computing platforms.
We can conclude that sequential algorithms with memory intensive operations like Liblinear can perform very well if datasets can fit in memory.
For massive datasets, if limited by machine resources, Mahout with its online algorithm is a choice with relatively low precision.
If we have enough machines, as Spark outperforms Hadoop for LR model training, we recommend to choose between parallel sublinear method and parallel gradient descent method (both implemented in Spark) to trade off between running time and precision.
Our future work will focus on combining algorithm design and computing platforms to optimize classifiers in machine learning.
